### 2019-10-15 ch1_introduction/content.tex
reshuffle

\begin{itemize}
	\item We \textbf{develop benchmarking strategies} for assessing the performance of computational tools constrained by low availability of novel types of real single-cell data (Chapter~\ref{chap:dyngen}). \textit{In silico} simulations of individual cells are used to help kick-start emerging domains much more safely and allow anticipation of future technological developments by already developing computational tools.
	\item We apply this strategy to \textbf{perform a comparison of TI methods} (Chapter~\ref{chap:dynbenchmark}). Trajectory inference is one of the largest categories of all the novel single-cell omics tools, yet a comprehensive and quantitative study of the advantages and disadvantages of the numerous tools was hitherto lacking. We provide a set of guidelines for end-users wishing to infer trajectories. Our pipeline, datasets, metrics, and containerised wrappers of TI methods are made publicly available for developers to use.
	\item We develop \textbf{a toolkit to infer, visualise and interpret single-cell trajectories} using more than 50 different TI methods (Chapter~\ref{chap:dyno}). dyno provides downstream analysis such as: visualising a trajectory in a low-dimensional space or a heatmap, detecting genes differentially expressed at different stages of the tragectory, comparing multiple trajectories in a common dimensionality reduction, and manipulating the trajectory (e.g. adding directionality or adding annotation).
	\item We introduce a TI method specialised in \textbf{inferring linear trajectories} (Chapter~\ref{chap:scorpius}). Despite linear TI being the most simple but commonly used form of trajectory inference, the benchmark demonstrated that most TI methods are not capable of producing accurate models of linear datasets.
	\item We develop an algorithm for  \textbf{inferring gene regulatory interactions at a single-cell level} (Chapter~\ref{chap:bred}). We demonstrate our NI methodology by inferring regulatory interactions for 25'000 hematopoietic cells, as well as for 15'000 cancer-related bulk profiles. % TODO: was this chapter written in the end?
	\item We provide a tool for  \textbf{analysing the topological properties of large, evolving networks} and use this to iteratively optimise GRN predictions (Chapter~\ref{chap:incgraph}).
	\item We reflect on \textbf{challenges in performing self-assessments of TI methods} (Chapter~\ref{chap:selfassessment}).
	Reproducability problems of TI methods are partially due to low rates of quantitative self-assessment. We look at 75 manuscripts of TI methods and identify key challenges the authors had in benchmarking their tool. For each of these challenges, we provide a summary of how we and others have successfully tackled these challenges, in order to spur developers to perform more self-assessments.
	\item Finally, we \textbf{summarise our experience in benchmarking computational methods} by providing a list of essential guidelines in benchmarking computational tools (Chapter~\ref{chap:guidelines}).
\end{itemize}

\section{List of contributions}

\subsection{First-author publications}

>>>>

\begin{itemize}
	\item We \textbf{develop benchmarking strategies} for assessing the performance of computational tools constrained by low availability of novel types of real single-cell data (Chapter~\ref{chap:dyngen}). \textit{In silico} simulations of individual cells are used to help kick-start emerging domains much more safely and allow anticipation of future technological developments by already developing computational tools.
	\item We apply this strategy to \textbf{perform a comparison of TI methods} (Chapter~\ref{chap:dynbenchmark}). Trajectory inference is one of the largest categories of all the novel single-cell omics tools, yet a comprehensive and quantitative study of the advantages and disadvantages of the numerous tools was hitherto lacking. We provide a set of guidelines for end-users wishing to infer trajectories. Our pipeline, datasets, metrics, and containerised wrappers of TI methods are made publicly available for developers to use.
	\item We develop \textbf{a toolkit to infer, visualise and interpret single-cell trajectories} using more than 50 different TI methods (Chapter~\ref{chap:dyno}). dyno provides downstream analysis such as: visualising a trajectory in a low-dimensional space or a heatmap, detecting genes differentially expressed at different stages of the tragectory, comparing multiple trajectories in a common dimensionality reduction, and manipulating the trajectory (e.g. adding directionality or adding annotation).
	\item We introduce a TI method specialised in \textbf{inferring linear trajectories} (Chapter~\ref{chap:scorpius}). Despite linear TI being the most simple but commonly used form of trajectory inference, the benchmark demonstrated that most TI methods are not capable of producing accurate models of linear datasets.
	\item We develop an algorithm for  \textbf{inferring gene regulatory interactions at a single-cell level} (Chapter~\ref{chap:bred}). We demonstrate our NI methodology by inferring regulatory interactions for 22'122 hematopoietic cells, as well as for 14'963 bulk profiles of cancer tumours.
	\item We provide a tool for  \textbf{analysing the topological properties of large, evolving networks} and use this to iteratively optimise GRN predictions (Chapter~\ref{chap:incgraph}).
	\item We reflect on \textbf{challenges in performing self-assessments of TI methods} (Chapter~\ref{chap:selfassessment}).
	Reproducability problems of TI methods are partially due to low rates of quantitative self-assessment. We look at 75 manuscripts of TI methods and identify key challenges the authors had in benchmarking their tool. For each of these challenges, we provide a summary of how we and others have successfully tackled these challenges, in order to spur developers to perform more self-assessments.
	\item Finally, we \textbf{summarise our experience in benchmarking computational methods} by providing a list of essential guidelines in benchmarking computational tools (Chapter~\ref{chap:guidelines}).
\end{itemize}

\section{List of contributions}
\subsection{First-author publications}
